Lifting CDCL to Numerical Domains

### 2016-10-07 contributions.tex
decision text

\section{Contributions}\label{contribution}
The goal of this paper is to instantiate ACDCL with 
different relational and non-relational numerical abstract 
domains and evaluate the performance of our solver with 
propositional-level solvers. Experimental evaluation on different
benchmarks shows that ACDCL is able to suitably exploit the 
expressivity of richer abstract domains as well as precision 
of CDCL architecture. This is attributed to the intelligent decision
heuristics which exploits the high-level structure of the problem 
combined with stronger deductions and clause learning mechanism aided 
by the richer abstract domains.

>>>>

\section{Contributions}
In this paper, we make the following contributions.
\begin{enumerate}
\item We present a novel program analysis technique based on 
abstract conflict driven clause learning algorithm~\cite{dhk2013-popl}.
To this end, we present a procedure for lifting CDCL algorithm to 
richer lattice structure such as {\em interval} and {\em octagon} 
abstract domains.

\item To enable lifting of CDCL algorithm, we implement 
our own template-based numerical abstract domains that 
supports complementation property.  Our proposed technique 
is embodied in our tool called, {\em Titlis}.  To the best 
of our knowledge, this is the first instantiation of CDCL 
algorithm with relational abstract domains.

\item  Experiments on a set of benchmarks drawn from bit-vector 
regression suite in sv-comp, numerical programs and few 
control-intensive benchmarks from hardware domain shows that 
the strict generalization of CDCL algorithm effectively combines 
the efficiency of abstract domains while being path-sensitive and 
bit-precise at the same time.  This is attributed to the intelligent 
decision heuristics which exploits the high-level structure of the 
problem combined with stronger deductions and clause learning 
mechanism aided by the richer abstract domains.
\end{enumerate}

### 2016-10-07 decision.tex
decision text

\section{Decision}
A decision $\mathcal{D}$ is a meet irreducible that refines the 
current partial assignment.  A decision can be of the form 
$\langle x \preceq c \rangle$ or $\langle x \succeq c \rangle$, 
where $c$ is the bound and $x$ is a program variable.  A 
decision $\mathcal{D}$ must always be consistent with respect 
to the trail $\mathcal{T}$, that is, it must not contradict with 
the elements in the trail.  Each variable in propositional CDCL 
is assigned at most once at a particular decision level.  However, 
a variable can be assigned several times in our lifting of CDCL to 
numerical domains, each time with increasingly precise bounds.

@@

We implemented several decision heuristics namely, {\em Berkmin}~\cite{}, 
{\em longest-range}, {\em ordered}, {\em random}.  The {\em ordered} 
decision heuristics creates an ordering among non-singleton variables, 
making decision on conditional variables (variables that appear in 
conditional branches) first before choosing numerical variables.  Thus, 
this heuristics is path-sensitive which gives significant speed-up for 
control-intensive benchmarks.

>>>>

\section{Decision}
A decision $\mathcal{D}$ is a meet irreducible that refines the 
current partial assignment.  A decision can be of the form 
$\langle x \preceq c \rangle$ or $\langle x \succeq c \rangle$, 
where $c$ is the bound and $x$ is a program variable.  A 
decision must always be consistent with respect to the 
trail $\mathcal{T}$, that is, it must not contradict with 
the elements in the trail.  Each variable in propositional CDCL 
is assigned at most once at a particular decision level.  However, 
a variable can be assigned several times in our lifting of CDCL to 
numerical domains, each time with increasingly precise bounds.

@@

{\em Titlis} support several decision heuristics namely, {\em ordered} 
{\em longest-range}, {\em random}, {\em Berkmin}~\cite{} decision heuristics.  
The {\em ordered} decision heuristics creates an ordering among non-singleton 
variables, thereby making decisions on conditional variables (variables that 
appear in conditional branches) first before choosing numerical variables.  
This variable ordering is path-sensitive as it gives an effect of trace
partitioning.  A {\em longest-range} heuristics simply keeps track of the
interval $[l,u]$ range ($r=u-l$) of a variable and chooses a variable with 
the longest range value.  This ensures a fairness policy in selecting a 
variable since it guarantees that the intervals of variables are uniformly 
restricted.  A {\em random} decision heuristics arbitrarily picks a variable 
for making decision.  Whereas, a {\em berkmin} decision heuristics is inspired 
from decision heuristics used in Berkmin~\cite{eugoldberg07} SAT solver, which 
keeps track of the activity of a variable that participate in conflict clauses 
as well as variables that actively contribute to conflicts but do not explicitly 
appear in conflict clauses.  The set of conflict clauses is organized 
chronologically with the top clause as the one deduced in the last.  A 
branching variable is chosen among the free variables whose literals are 
in the top unsatisfied conflict clause.  A similar decision heuristics is 
also implemented in Chaff~\cite{chaff} SAT solver, that computes the activity 
of a variable as the number of occurrences of that variable in conflict 
clauses only.

A bound of a variable is heuristically chosen to be an approximation of the 
arithmetic average of the current bounds.  However, the polarity ($\preceq$ or
$\succeq$) of a variable is chosen randomly.
